---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyr)
library(tidyverse)
library(janitor)
```

Read the cleaned data into R.

```{r}
clean_mete <- read_csv("Clean data /meteorite_cleaned.csv")
```

Find the names and years found for the 10 largest meteorites in the data.

```{r}
clean_mete %>% 
  select(c("name", "year", "mass_g"))%>%
  slice_max(mass_g, n= 10, with_ties = FALSE )
#this list returns top 10 with year and mass.  
```

Find the average mass of meteorites that were recorded falling, vs. those which were just found.
```{r}
#3.5 Writing function/program to perform statistical analysis
clean_mete %>% 
  group_by(fall) %>% 
  summarise(average = mean(mass_g, na.rm = TRUE))
#on average, meteorites fell was a lot bigger than the ones are found.
```

Find the number of meteorites in each year, for every year since 2000.
```{r}
clean_mete %>% 
  filter(year >= 2000) %>%
  group_by(year) %>% 
  summarise(count = n())
#from 2000 to 2003 there were continuous increasing every year.
#in general there aren't any clear trend. 
```

Write a function called get_meteorite_count() that takes in two arguments: a dataframe and a year, and returns the phrase “In (input_year), there were (number of meteorites that year) meteorites.” IF there are 1 or more meteorites that year. If there are no meteorites in the year, return the phrase “As far as we know, there were no meteorites that year”.

```{r}
get_meteorite_count<- function(d,y){
  
  num = d %>% 
  filter(year == y) %>%
  summarise(count = n()) %>% 
  pull()
  
  if(num >= 1){
    return(paste0("In", y, ", there are ", num," meteorites."))
  }else{
    if(num < 1){
    return("As far as we know,there were no meteorites that year")
  }
  }
}
```

Write down ideas for how the function you’ve written could be improved and how you could test that the function works as it should. Please keep these ‘test plans’ in the final .Rmd, especially if you are completing the PDA.
```{r}
#My function didn't think about NA from the data sets.
#It will let non_numeric input of years pass.

#I can test with some data frame contains NA,
#Test year input as strings and logical.

```

Write a unit test for your function to check whether non-numeric inputs (character and logical) for the year argument return an error.
```{r}
#4.1 Test plans 
library(testthat)
test_that("Non-numeric input returns an error", {
  expect_error(get_meteorite_count(clean_mete, "yes"))
  expect_error(get_meteorite_count(clean_mete, TRUE))
})
#My old function didn't pass the test.
```

Improve your function based on the result of your test. Instead of going back and changing the function, please create a new function that is the improved version called refactor_get_meteorite_count(). Test your function to make sure it now gives an error when you input a logical or character value as the year. Hint: You can use the stop function to create your own error message.
```{r}
#4.6 Performing test and revising code
library(assertr)
refactor_get_meteorite_count<- function(d,y){
  stopifnot(is.numeric(y))
  
  num = d %>% 
  filter(year == y) %>%
  summarise(count = n()) %>% 
  pull()
  
  if(num >= 1){
    return(paste0("In", y, ", there are ", num," meteorites."))
  }else{
    if(num < 1){
    return("As far as we know,there were no meteorites that year")
  }
  }
}
```

```{r}
test_that("Non-numeric input returns an error", {
  expect_error(refactor_get_meteorite_count(clean_mete, "yes"))
  expect_error(refactor_get_meteorite_count(clean_mete, TRUE))
})
```


3.6 Writing function/program to create visualisations

```{r}
clean_mete %>% 
  group_by(fall) %>% 
  summarise(average = mean(mass_g, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x = fall, y = average/1000)+
  geom_col(aes(fill = fall))+
  labs(x = "\nFall or Found",
       y = "Mass(kg)",
       fill = "",
       title = "Weight Compareing of Fall vs Found\n")
```
#this plot links back to one of the analysis question,(Find the average mass of meteorites that were recorded falling, vs. those which were just found.)
#It shows that on average, meteorites fell was a lot smaller than the ones are found.


week 11 
Based on the mass and year, create a decision tree model of your meteorites data which predicts whether a meteorite was seen falling, or was discovered after its impact (fell/found).

First, you want to convert any character column to a factor. Next (and most importantly) split your dataset into training and testing datasets. Then create a decision tree with rpart and plot it using rpart.plot. Finally, test and add your predictions to your data, and create a confusion matrix.

```{r}
library(rpart)
library(rpart.plot)
library(dplyr)
```

```{r}
clean_mete %>% head()
mete_tree <- clean_mete %>% select(mass_g, year,fall) %>% 
  mutate(fall = factor(fall, levels = c("Found", "Fell"), labels = c("Found", "Fell"))) %>% 
   na.omit()
mete_tree
```
```{r}
set.seed(19)

n_data <- nrow(mete_tree)

test_index <- sample(1:n_data, size = n_data*0.2)

# create test set
test  <- slice(mete_tree, test_index)

# create training set
train <- slice(mete_tree, -test_index)
```

```{r}
test %>%
 janitor::tabyl(fall)
train %>%
 janitor::tabyl(fall)
```

```{r}
mete_fit <- rpart(
  formula = fall ~ ., 
  data = train, 
  method = 'class'
)

rpart.plot(mete_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4)
```
```
Taking all of our data, there are 15.47% likely to be fell.

If the year is 1936 or later, which is 78.89% of the data, there are 7.9% chance to be fell.

If the year is before 1936  and mass is greater than 613,000g, (1.01% of the data), are less than 0.001% likely to be fell. 

If the year is before 1836, mass between 10,800g and 613,000g(8.85% of the data) are 28.45% likely to  be fell.

If the year is before 1838 and the mass is between 55,500g and 613,000g (0.34% of the data), it's 23.08% likely to be fell.

If the year is earlier than 1838, and weight is less than 55,500g (57% of the data), are 95.45% likely to be fell. 

If the year is between 1884 and 1919, and the mass is between 1313 g and 10,800g, (3.12% of the data), it's 35% likely to be fell. 

If the year is between 1919 and 1936, the mass is between 1313 and 10,800g (2.23% of the data), are 55.81% likely to fell. 

If the year is between 1884 and 1936, and mass is less than 1313g,(0.78% of the data) are 70% likely to be fell.

If the year is before 1884,mass less than 10,800g(4.21% of the data), are 76.54% likely to be fell
```
```{r}
library(modelr)

# add the predictions
mete_test_pred <- test %>%
  add_predictions(mete_fit, type = 'class')
```

```{r}
library(yardstick)
conf_mat <- mete_test_pred %>%
              conf_mat(truth = fall, estimate = pred)

conf_mat
```

```
This model performs quite good, it's mostly correct. 799 true possitive and 43 true negative 
with 120 errors. 
```

```{r}
accuracy <- mete_test_pred %>%
 accuracy(truth = fall, estimate = pred)

accuracy 
```
```
It's 87.53% correct 
```
