---
title: "Penguins PDA"
output: html_notebook
---


# 1. Exploratory analysis
Explore the dataset using visualisations, and keep in mind the question you’re trying to answer above.

Write a short summary of the results.
```{r}
library(palmerpenguins)
library(tidyverse)
library(e1071)
library(infer)
penguins <- palmerpenguins::penguins 
```

```{r}
summary(penguins)
penguins %>% 
  ggplot(aes(species,bill_length_mm))+
  geom_boxplot(aes(colour = species))+
  geom_jitter(aes(colour = species), alpha = 0.5)
```
```
From the graph above, chinstrap has the longest bill and 
adelie has the shortest. gentoo and chinstrap looks very similar. with in each 
species, the distribution seems normal.
```

# 2. Summary statistics
Calculate the following summary statistics to describe bill length for each species:

measures of centrality (mean & median)
spread (standard deviation)
skewness (you can use the skewness() function from the e1071 package)
Write a short summary of the results.

## 4.4 Summary statistics including measures of centrality, spread and skewness
```{r}
penguins %>% 
  group_by(species) %>% 
  summarise(mean_bill = mean(bill_length_mm, na.rm = TRUE),
            median = median(bill_length_mm, na.rm = TRUE),
            skewness = skewness(bill_length_mm, na.rm = TRUE, type = 1)
            )
```
```
From the table above, the average bill_lenght matachs with the box_plot(Chinstrap has the longest then gentoo and adelie),
adelie and gentoo, chinstrp and adelie are symmtric,
and gentoo are moderately right skewed 
```


# 3. Hypothesis test & probability
Perform a suitable hypothesis test to check whether the mean bill length of Gentoo is significantly longer than the mean bill length of Adelie.

Visualise the p-value on the null distribution.
Write an interpretation of your findings.


## 4.1 Exploratory data analysis and targeted data analysis (hypothesis testing)
## 4.3 Probability and probability distributions

My Null(H0) and alternative(H1) hypothesis as below, my alpha level is 0.05
$$
N0: \mu_{gentoo\ bill\ length} = \mu_{adelie\ bill\ length} \\
N1: \mu_{gentoo\ bill\ length} > \mu_{adelie\ bill\ length}
$$


```{r}
null_distribution <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo")) %>% 
  specify(bill_length_mm ~ species) %>% 
  hypothesise(null = 'independence') %>% 
  generate(rep = 1000, type = 'permute') %>% 
  calculate(stat = "diff in means", order = c('Gentoo', 'Adelie'))

null_distribution
```



```{r}
observed_stats <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo")) %>% 
  specify(bill_length_mm ~ species) %>% 
  calculate(stat = "diff in means", order = c('Gentoo', 'Adelie'))

observed_stats 
```
```{r}
null_distribution %>% 
  visualise()+
  shade_p_value(obs_stat = observed_stats, direction = "right")
```
```{r}
p_value <- null_distribution %>% 
  get_p_value(obs_stat = observed_stats, direction = "right")
p_value
```
```
p_value is how likely it is to see a result as extreme as observed result, if the null hypothesus is true.
we got p_value < 0.0001, this means it's not very likely to see a dataset such as ours, if the HO is true. 
It is a lot less than our alpha level, 
the difference is significant and we reject H0 and 
found evidence to support H1. 
```

Create a correlation matrix using GGally::ggcorr() that shows the correlation between 
bill length, bill depth, flipper length, and body mass across species.
(Hint: you can add correlation coefficient labels to your plot if you look at the ?ggcorr help file)

Write a short summary of your results.

```{r}
penguins %>% 
  select("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g") %>% 
GGally::ggcorr(label = TRUE)
```
```
body mass and flipper length has a very strong possitive correlation, 
body mass also has strong possitive correlation with bill length, 
flipper length has a strong positive correlation with bill length, 
bill length has a week negtive correlation with bill depth. 
```
