---
title: "CitiBike Report"
output: html_notebook
---

## 1. Loading data and tidying up 
```{r}
library(tsibbledata)
library(tidyverse)
library(lubridate)
library(tsibble)
library(leaflet)
nyc_bikes_df <- nyc_bikes
```

```{r}
nyc_bikes_short <- nyc_bikes_df %>% 
  select(start_time, stop_time, birth_year, gender, type) %>% 
  mutate(year = year(start_time),
         month = month(start_time, label = TRUE),
         day = as.Date(start_time),
         wday  = wday(start_time,label = TRUE),
         week = week(start_time),
         time_ride = stop_time - start_time,
         age = as.numeric(year - birth_year))

bike_90 <-nyc_bikes_short%>%
  index_by(year) %>%
  summarise(p90 = quantile(time_ride, probs=0.9, na.rm=TRUE))

nyc_bikes_short_90 <- nyc_bikes_short %>% 
  mutate(time_ride = as.numeric(time_ride)) %>% 
  filter(time_ride < bike_90$p90)
#nyc_bikes_short doesn't have locations data 
#nyc_bikes_short_90 has 90% of nyc_bikes_short based on riding time
```
## 2 Overall pattern of the year

2.1 Daily distribution over 2018

```{r}
day_nyc <- nyc_bikes_short_90 %>% 
  group_by(day) %>% 
  count()

day_nyc %>% 
  ggplot(aes(day, n))+
  geom_line()+
  labs(x = "Days in 2018",
       y = "Number of Bike Hire",
       title = "Bike hire for each day in 2018\n")
```
From the grapg above, it suggestes that from Jan to Jul bike hiring increasing
and the number of hiring each day starts to drop from summer to winter months.

## 2.2 Monthly pattern over 2018
```{r}
month_nyc <- nyc_bikes_short_90 %>% 
  group_by(month) %>% 
  count()

month_nyc %>% 
  ggplot(aes(month, n))+
  geom_line(group = 1, size = 1, col = "lightblue")+
  geom_point(col = "steelblue")+
  labs(
    x="\nMonth",
    y = "Number of Bike Hire",
    title = "Number of bike hire for each month",
    subtitle = "Data from 2018 citibike NYC\n"
  )
```
From the monthly disribution, we can notice that the pattern of bike hiring is 
very similar to the daily graph and suggests summer is the peak of the year for
bike hiring.

## 2.3 weekly patterns 

```{r}
wday_nyc <- nyc_bikes_short_90%>% 
  group_by(wday) %>% 
  count()



wday_nyc %>% 
  ggplot(aes(wday, n))+
  geom_point(size = 2, col = "steelblue")+
  geom_line(group = 1, alpha = 0.7, col = "steelblue")+
   labs(
    x="Weekdays",
    y = "Number of Bike Hire",
    title = "Number of bike hire for each day of week",
    subtitle = "Data from 2018 citibike NYC\n"
  )
```
Across the whole year of 2018, the graph suggests that the numbers of bike hiring 
is higher during the weekdays and it lower for weekends. It could be that the 
most common usage are for commuting. 

# 3. Compare between genders, ages and types.
## 3.1 Gender overall and over the year 
```{r}
nyc_bikes_short_90 %>% 
  group_by(gender) %>% 
  count() %>% 
  ggplot(aes(gender, n))+
  geom_col(aes(fill = gender, col = gender))+
  labs(x = "Gender",
       y = "Number of bikr hire",
       title = "Gender bike hire in 2018")
```
This graph shows an overview of bike hiring for each gender in 2018. 
Male users are a lot higher than the other two groups. 

```{r}
gender_m <- nyc_bikes_short_90 %>% 
  group_by(gender) %>% 
  index_by(month) %>% 
  count()

gender_m %>% 
  ggplot(aes(x = month, y = n))+
  geom_line(aes(col = gender), group = 1)+
  geom_point(aes(col = gender), size = 2)+
  facet_wrap(~gender,ncol =1)+
  labs(x = "Month",
       y = "Number of bike hire",
       title = "Bike hire pattern for each gender")
```
The hiring pattern seems to be similar as we descoved before. All groups have
a peak around summer months and male hiring are a lot higher 
than the other two groups.

## 3.2 Age overall and over weekdays 

I used age groups are as following: 18-30, 30-60, 60 and above.
I am interested in mostly over 60 group as I think they will 
show different pattern over a week as youger groups. 

```{r}
nyc_bikes_short_90 %>% 
   mutate( age_group = case_when(age < 18 ~ "Under 18",
                        age < 30 ~"18 to 30",
                        age < 60 ~"30 to 60",
                        TRUE ~"60 and above")) %>% 
  group_by(age_group) %>% 
  count() %>% 
  ggplot(aes(age_group, n))+
  geom_col(aes(fill = age_group, col = age_group))+
  labs(x = "Age Group",
       y = "Number of bikr hire",
       fill = "Age Group",
       col = "Age Group",
       title = "Bike hiring for each age group")
```
The overall graph shows that most of the bike hiring are from 30-60 age
group. However, it is the biggest age group. It is not very fair to use 
this grouping. 


```{r}
nyc_bikes_short_90 %>% 
   mutate( age_group = case_when(age < 18 ~ "Under 18",
                        age < 30 ~"18 to 30",
                        age < 60 ~"30 to 60",
                        TRUE ~"60 and above")) %>% 
  group_by(age_group) %>% 
  index_by(wday) %>% 
  count() %>% 
  ggplot(aes(wday, n))+
  geom_point(size = 2, aes(col = age_group))+
  geom_line(group = 1, alpha = 0.7, aes(col = age_group))+
  facet_wrap(~age_group, ncol = 1,scales = "free_y")+
   labs(
    x="Weekdays",
    y = "Number of Bike Hire",
    col = "Age Group",
    fill = "Age Group",
    title = "Bike hiring for each age group of weekdays"
  )
```
Supprisingly, for all the groups it shows similar pattern as the whole data
group(bike hiring are mostly in weekdays), Even for over 60 groups which 
higher percentage of them among general population don't need to commute for
work. 

## 3.3 Types of customers, overall and over the year

```{r}
nyc_bikes_short_90 %>% 
  group_by(type) %>% 
  count() %>% 
  ggplot(aes(type, n))+
  geom_col(aes(fill = type, col = type))+
  labs(x = "Type of costomers",
       y = "Number of bikr hire",
       col = "Costomer types",
       fill = "Costomer types",
       title = "Bike hire for differnt costomer types")
```
Most the customers are subscribers and only small percentages are not.

```{r}
type_d <- nyc_bikes_short_90 %>% 
  group_by(type) %>% 
  index_by(day) %>% 
  count()

type_d %>% 
  ggplot(aes(x = day, y = n))+
  geom_line(aes(col = type))+
  facet_wrap(~type,ncol =1, scales = "free_y")+
  labs(x = "Day",
       y = "Number of bike hire",
       col = "Type",
       title = "Bike Hire Pattern for Different Types of Hiring")
```
This graph shows the patterns of bike hiring of differnt types over 2018.
for subscribers, they have very similar pattern as the overall data group. 
High peaks in summer and lower number of hiring in winter.
However for "Cusomer" type, the pattern shows sharp increasing and sharp drop
throughout the year and it also has time period that no "customer" are hiring 
in a day. This is due to the fact that not many users are this type and 
it not expected to see a steaby amount everyday. 

# 4. Locations 

```{r}
nyc_bikes_df %>% 
  group_by(start_long, start_lat) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(5) %>% 
leaflet() %>% 
  addTiles() %>% 
  addMarkers(
    lng = ~start_long,
    lat = ~start_lat,
    clusterOptions = markerClusterOptions()
  )
```
The spacial graph above shows the top 5 most popular stations,
all of they are very close to big public transport stations and most of them
are located in downtown. 


```{r}
nyc_bikes_df %>% 
  filter(type == "Customer") %>% 
leaflet() %>% 
  addTiles() %>% 
  addMarkers(
    lng = ~start_long,
    lat = ~start_lat,
    clusterOptions = markerClusterOptions()
  )
```
Assume that all the customer types are visitors, they should use the top 5 
stations to travel into the city. From the graph above they do seems to use the 
major stations more but there are small other stations been used. 


# 5 Data ethics 
Data used is open data from CitiBike and collected with consent 

Analysis only looks at trends rather than one spcific data 

This set had a sample of 10 bike so it's not very representative for whole 
population 


# 6 Conclusion and recommendations 
The KPI is to increase bike hiring, we noticed the difference in male and female
customers, the company should encourage female customers. Marketing among female
population.
Encourage visitors to hire citibike, we can have more bikes around big public stations. 